```{r}
title: "PSID: Predicting Income with Social Data"
output:
  html_document:
    df_print: paged
```
### PSID DATA

Source: Code Academy - Analyze Data with R

The Panel Survey of Income Data is the longest running longitudinal household survey in the world. Survey responses related to social, economic, and health outcomes have been collected from the families and their descendants since 1968. This dataset is widely used by social scientists to investigate the relationship between individual characteristics, like gender or age, and broader socioeconomic outcomes like education achievement and lifetime income.

This project will use PSID data and linear regression to predict the labor-derived income of survey respondents based on the following set of variables:

gender: the gender, female-identifying, male-identifying, or other, of a respondent
age: the age of the respondent
married: the marital status, unmarried, married, or divorced, of a respondent
employed: the employment status of the respondent at the time of survey collection
educated_in_us: whether the respondent went to primary school in the United Statues
highest_degree: the highest educational degree obtained by the respondent
education_years: the total number of years of formal education completed by the respondent
labor_income: the yearly income earned by the respondent from a salary or hourly employment

###Investigate the data
```{r}
# Load packages
library(ggplot2)
library(dplyr)
library(modelr)
```

```{r}
# Load the data set
psid <- read.delim("psid.txt", header = TRUE, sep = ",")
```


```{r}
# View data structure
str(psid)
head(psid)
```


```{r}
# Plot age
age_dist <- psid %>% ggplot(aes(age)) + geom_bar()
age_dist
```


```{r}
# Filter to reasonable age group
psid_clean <- psid %>% filter(age>18,age<75)
psid_clean
```


```{r}
# Plot filtered age
psid_clean_viz <- psid_clean %>% ggplot(aes(age)) + geom_bar()
psid_clean_viz
```


```{r}
# Plot education
educ_dist <- psid_clean %>% ggplot(aes(education_years,education_years)) + geom_boxplot()
educ_dist
```


```{r}
# filter to reasonable education levels
educ_clean <- psid_clean %>% filter(education_years>5,education_years<25)
educ_clean
```


```{r}
# plot income
labor_income_dist <- psid_clean %>% ggplot(aes(labor_income,labor_income)) + geom_boxplot()
labor_income_dist
```


```{r}
# View income summary
summary(psid_clean$labor_income)
```


```{r}
# plot mean income by age
avg_income_age <- psid_clean %>%
                 group_by(age) %>%
                 summarise(mean_income=mean(labor_income, na.rm = TRUE)) %>%
                ggplot(aes(age,mean_income)) + geom_point()
avg_income_age
```

### Build model and assess fit

```{r}
# subset data points into train and test sets
set.seed(123)
sample <- sample(c(TRUE, FALSE), nrow(psid_clean), replace = T, prob = c(0.6,0.4))
```


```{r}
# Define train and test set
train <- psid_clean[sample, ]
test <- psid_clean[!sample, ]
```


```{r}
# build model
model <- lm(labor_income ~ education_years, data = train)
```


```{r}
# Plot LOESS model
plot <- train %>% ggplot(aes(education_years,labor_income)) +
        geom_point() +
        geom_smooth(method = "lm") +
        geom_smooth(se = FALSE, color = "red")
plot
```


```{r}
# Compute r-squared
r_sq <- summary(model)$r.squared * 100
r_sq
```


```{r}
sprintf("Based on a simple linear regression model, we have determined that %s percent of the variation in respondent income can be predicted by a respondent's education level.", r_sq)
```

### Build comparison model and analyze results

```{r}
# build second model
model_2 <- lm(labor_income ~ education_years + age + gender, data = train)
r_sq_2 <- summary(model_2)$r.squared * 100
r_sq_2
```


```{r}
sprintf("Based on a simple linear regression model, we have determined that %s percent of the variation in respondent income can be predicted by a respondent's education level, age and gender.", r_sq_2)
```


```{r}
# plot predictions versus observed
plot <- test %>%
        add_predictions(model_2) %>%
  ggplot(aes(age, labor_income)) +
  geom_point() +
  geom_line(aes(y = pred), color = "blue") 

plot
```


```{r}
# write out model results
summary(model_2)
```


```{r}

```


```{r}

```

